---
title: "10_controlmodels"
author: "Chris Cox"
date: "2025-01-06"
output: pdf_document
editor_options: 
  chunk_output_type: console
---

# No Discontinuity Assumption

```{r, eval = FALSE}
# Change preprocessing to allow acceleration to persist over days:
d_multivariate_no_dis <- d_multivariate %>%
  group_by(CoupleID) %>%
  arrange(CoupleID, time) %>%
    mutate(
      across(
        c(MotherArousal_s, MotherValence_s, 
          FatherArousal_s, FatherValence_s),
        list(
          Prev = ~lag(.),
          PrevPrev = ~lag(., 2),
          PrevPrevPrev = ~lag(., 3),
          PrevPrevPrevPrev = ~lag(., 4),
          PrevPrevPrevPrevPrev = ~lag(., 5)
        )
      )
    ) %>%
  group_by(CoupleID, Leave) %>%
  arrange(CoupleID, Leave, time) %>%
  mutate(
    FatherValenceVel = (FatherValence_s - FatherValence_s_Prev),
    MotherValenceVel = (MotherValence_s - MotherValence_s_Prev),
    FatherValenceAcc = (FatherValence_s - 2*FatherValence_s_Prev + FatherValence_s_PrevPrev),
    MotherValenceAcc = (MotherValence_s - 2*MotherValence_s_Prev + MotherValence_s_PrevPrev)
  ) %>%
  ungroup()

CoupledOscillatorModel_no_discontinuity <- brm(
            f_coupled_dynamics_acc,
            data = d_multivariate_no_dis,
            family = gaussian(),
            prior = p_coupled_dynamics, 
            sample_prior = "yes",
            chains = 2,
            cores = 64,
            warmup = 500,
            iter = 2000,
            control = list(adapt_delta = 0.90, max_treedepth = 10),
            backend = 'cmdstan')
```

## Extract Estimates for Supplementary Materials

```{r}
#CoupledOscillatorModel_no_discontinuity <- readRDS(here("FinalModels", "CoupledOscillatorModel_no_discontinuity.rds"))
tests_no_discontinuity <- extract_all_tests(CoupledOscillatorModel_no_discontinuity)
write.csv(tests_no_discontinuity, here("StatsForBodyText", "coupledoscillator_tests_no_discontinuity.csv"), row.names = FALSE)
```

# Carry Non-Missing Values Forward

```{r, eval = FALSE}
# First, let's handle the NA values within each StudyDay
d_multivariate_CarryForwardNonNAs <- d_multivariate %>%
  group_by(CoupleID, StudyDay) %>%
  arrange(CoupleID, StudyDay, time) %>%
  mutate(
    # For Father's valence, carry forward the last non-NA value within StudyDay
    FatherValence_s = if_else(
      is.na(FatherValence_s) & 
      !is.na(lag(FatherValence_s)),
      lag(FatherValence_s),
      FatherValence_s
    ),
    # For Mother's valence, carry forward the last non-NA value within StudyDay
    MotherValence_s = if_else(
      is.na(MotherValence_s) & 
      !is.na(lag(MotherValence_s)),
      lag(MotherValence_s),
      MotherValence_s
    )
  ) %>%
  # Now proceed with the rest of preprocessing
  mutate(
    across(
      c(MotherArousal_s, MotherValence_s, 
        FatherArousal_s, FatherValence_s),
      list(
        Prev = ~lag(.),
        PrevPrev = ~lag(., 2),
        PrevPrevPrev = ~lag(., 3),
        PrevPrevPrevPrev = ~lag(., 4),
        PrevPrevPrevPrevPrev = ~lag(., 5)
      )
    )
  ) %>%
  ungroup() %>%
  group_by(CoupleID, Leave) %>%
  arrange(CoupleID, Leave, time) %>%
  mutate(
    FatherValenceVel = (FatherValence_s - FatherValence_s_Prev),
    MotherValenceVel = (MotherValence_s - MotherValence_s_Prev),
    FatherValenceAcc = (FatherValence_s - 2*FatherValence_s_Prev + FatherValence_s_PrevPrev),
    MotherValenceAcc = (MotherValence_s - 2*MotherValence_s_Prev + MotherValence_s_PrevPrev)
  ) %>%
  ungroup()

f_coupled_dynamics_acc <- mvbf(
    # Valence position
    FatherValence_s ~ 0 + Leave + (Leave | CoupleID),
    MotherValence_s ~ 0 + Leave + (Leave | CoupleID),
    
    # Arousal position
    FatherArousal_s ~ 0 + Leave + (Leave | CoupleID),
    MotherArousal_s ~ 0 + Leave + (Leave | CoupleID),
    
    # Velocity equations without coupling
    FatherValenceVel ~ 0 + Leave + (Leave | CoupleID),
    MotherValenceVel ~ 0 + Leave + (Leave | CoupleID),
    
    # Acceleration equations with coupling
    FatherValenceAcc ~ 0 + Leave +
                       Leave:FatherValence_s + Leave:FatherValenceVel +
                       Leave:MotherValence_s:isProximate + Leave:MotherValenceVel:isProximate +
                       Leave:FatherValence_s:FatherArousal_s + Leave:FatherValenceVel:FatherArousal_s +
                       Leave:MotherValence_s:isProximate:MotherArousal_s + Leave:MotherValenceVel:isProximate:MotherArousal_s +
                      (Leave:FatherValence_s + Leave:FatherValenceVel + Leave:MotherValence_s:isProximate + Leave:MotherValenceVel:isProximate + Leave:FatherValence_s:FatherArousal_s + Leave:FatherValenceVel:FatherArousal_s + Leave:MotherValence_s:isProximate:MotherArousal_s + Leave:MotherValenceVel:isProximate:MotherArousal_s | CoupleID / StudyDay),
    
    MotherValenceAcc ~ 0 + Leave +
                       Leave:MotherValence_s + Leave:MotherValenceVel +
                       Leave:FatherValence_s:isProximate + Leave:FatherValenceVel:isProximate +
                       Leave:MotherValence_s:MotherArousal_s + Leave:MotherValenceVel:MotherArousal_s +
                       Leave:FatherValence_s:isProximate:FatherArousal_s + Leave:FatherValenceVel:isProximate:FatherArousal_s +
                      (Leave:MotherValence_s + Leave:MotherValenceVel + Leave:FatherValence_s:isProximate + Leave:FatherValenceVel:isProximate + Leave:MotherValence_s:MotherArousal_s + Leave:MotherValenceVel:MotherArousal_s + Leave:FatherValence_s:isProximate:FatherArousal_s + Leave:FatherValenceVel:isProximate:FatherArousal_s | CoupleID / StudyDay))

p_coupled_dynamics <- c(
  prior(normal(0, 2), class = "b"),
  prior(normal(1, 2), class = "sd", resp = "FatherValenceAcc"),
  prior(normal(1, 2), class = "sd", resp = "FatherValences"),
  prior(normal(1, 2), class = "sd", resp = "FatherValenceVel"),
  prior(normal(1, 2), class = "sd", resp = "MotherValenceAcc"),
  prior(normal(1, 2), class = "sd", resp = "MotherValences"),
  prior(normal(1, 2), class = "sd", resp = "MotherValenceVel"),
  prior(normal(1, 2), class = "sigma", resp = "FatherValenceAcc"),
  prior(normal(1, 2), class = "sigma", resp = "FatherValences"),
  prior(normal(1, 2), class = "sigma", resp = "FatherValenceVel"),
  prior(normal(1, 2), class = "sigma", resp = "MotherValenceAcc"),
  prior(normal(1, 2), class = "sigma", resp = "MotherValences"),
  prior(normal(1, 2), class = "sigma", resp = "MotherValenceVel"),
  prior(lkj(2), class = "rescor")
)

CoupledOscillatorModel_CarryForwardNonNAs <- brm(
            f_coupled_dynamics_acc,
            data = d_multivariate_CarryForwardNonNAs,
            family = gaussian(),
            prior = p_coupled_dynamics, 
            sample_prior = "yes",
            chains = 2,
            cores = 64,
            warmup = 500,
            iter = 2000,
            control = list(adapt_delta = 0.90, max_treedepth = 10),
            backend = 'cmdstan')
```

## Extract Estimates for Supplementary Materials

```{r}
#CoupledOscillatorModel_CarryForwardNonNAs <- readRDS(here("FinalModels", "CoupledOscillatorModel_CarryForwardNonNAs.rds"))
tests_CFNA <- extract_all_tests(CoupledOscillatorModel_CarryForwardNonNAs)
write.csv(tests_CFNA, here("StatsForBodyText", "coupledoscillator_tests_CFNA.csv"), row.names = FALSE)
```

# Kalman imputation

```{r, eval = FALSE}
d_multivariate_imputed <- d_multivariate %>%
  group_by(CoupleID) %>%
  arrange(time) %>%
  mutate(
    MotherValence_s = safe_kalman(MotherValence_s),
    FatherValence_s = safe_kalman(FatherValence_s),
    MotherArousal_s = safe_kalman(MotherArousal_s),
    FatherArousal_s = safe_kalman(FatherArousal_s)
  ) %>%
  ungroup() %>%
  mutate(
    across(
      c(MotherArousal_s, MotherValence_s, 
        FatherArousal_s, FatherValence_s),
      list(
        Prev = ~lag(.),
        PrevPrev = ~lag(., 2),
        PrevPrevPrev = ~lag(., 3),
        PrevPrevPrevPrev = ~lag(., 4),
        PrevPrevPrevPrevPrev = ~lag(., 5)
      )
    )
  ) %>%
  ungroup() %>%
  arrange(CoupleID, Leave, time) %>%
  group_by(CoupleID, Leave) %>%
  arrange(CoupleID, Leave, time) %>%
  mutate(
    FatherValenceVel = (FatherValence_s - FatherValence_s_Prev),
    MotherValenceVel = (MotherValence_s - MotherValence_s_Prev),
    FatherValenceAcc = (FatherValence_s - 2*FatherValence_s_Prev + FatherValence_s_PrevPrev),
    MotherValenceAcc = (MotherValence_s - 2*MotherValence_s_Prev + MotherValence_s_PrevPrev)
  ) %>%
  ungroup()
  
CoupledOscillatorModel_kalman <- brm(
            f_coupled_dynamics_acc,
            data = d_multivariate_imputed,
            family = gaussian(),
            prior = p_coupled_dynamics, 
            sample_prior = "yes",
            chains = 2,
            cores = 64,
            warmup = 500,
            iter = 2000,
            control = list(adapt_delta = 0.90, max_treedepth = 10),
            backend = 'cmdstan')
```

## Extract Estimates for Supplementary Materials

```{r}
#CoupledOscillatorModel_kalman <- readRDS(here("FinalModels", "CoupledOscillatorModel_kalman.rds"))
tests_kalman <- extract_all_tests(CoupledOscillatorModel_kalman)
write.csv(tests_kalman, here("StatsForBodyText", "coupledoscillator_tests_kalman.csv"), row.names = FALSE)
```

# Controlling for Time Difference
```{r, eval = FALSE}
f_coupled_dynamics_acc_time <- mvbf(
    FatherValence_s ~ 0 + Leave + (Leave | CoupleID),
    MotherValence_s ~ 0 + Leave + (Leave | CoupleID),
    FatherArousal_s ~ 0 + Leave + (Leave | CoupleID),
    MotherArousal_s ~ 0 + Leave + (Leave | CoupleID),
    FatherValenceVel ~ 0 + Leave + (Leave | CoupleID),
    MotherValenceVel ~ 0 + Leave + (Leave | CoupleID),
    
    # Modified acceleration equations with time difference interactions
    FatherValenceAcc ~ 0 + Leave +
                       Leave:FatherValence_s + Leave:FatherValenceVel +
                       Leave:MotherValence_s:isProximate + Leave:MotherValenceVel:isProximate +
                       Leave:MotherValence_s:isProximate:TimeDifference_s + Leave:MotherValenceVel:isProximate:TimeDifference_s +
                       Leave:FatherValence_s:FatherArousal_s + Leave:FatherValenceVel:FatherArousal_s +
                       Leave:MotherValence_s:isProximate:MotherArousal_s + Leave:MotherValenceVel:isProximate:MotherArousal_s +
                       Leave:MotherValence_s:isProximate:MotherArousal_s:TimeDifference_s + Leave:MotherValenceVel:isProximate:MotherArousal_s:TimeDifference_s +
                      (Leave:FatherValence_s + Leave:FatherValenceVel + 
                       Leave:MotherValence_s:isProximate + Leave:MotherValenceVel:isProximate +
                       Leave:MotherValence_s:isProximate:TimeDifference_s + Leave:MotherValenceVel:isProximate:TimeDifference_s +
                       Leave:FatherValence_s:FatherArousal_s + Leave:FatherValenceVel:FatherArousal_s +
                       Leave:MotherValence_s:isProximate:MotherArousal_s + Leave:MotherValenceVel:isProximate:MotherArousal_s +
                       Leave:MotherValence_s:isProximate:MotherArousal_s:TimeDifference_s + Leave:MotherValenceVel:isProximate:MotherArousal_s:TimeDifference_s | CoupleID / StudyDay),
    
    MotherValenceAcc ~ 0 + Leave +
                       Leave:MotherValence_s + Leave:MotherValenceVel +
                       Leave:FatherValence_s:isProximate + Leave:FatherValenceVel:isProximate +
                       Leave:FatherValence_s:isProximate:TimeDifference_s + Leave:FatherValenceVel:isProximate:TimeDifference_s +
                       Leave:MotherValence_s:MotherArousal_s + Leave:MotherValenceVel:MotherArousal_s +
                       Leave:FatherValence_s:isProximate:FatherArousal_s + Leave:FatherValenceVel:isProximate:FatherArousal_s +
                       Leave:FatherValence_s:isProximate:FatherArousal_s:TimeDifference_s + Leave:FatherValenceVel:isProximate:FatherArousal_s:TimeDifference_s +
                      (Leave:MotherValence_s + Leave:MotherValenceVel + 
                       Leave:FatherValence_s:isProximate + Leave:FatherValenceVel:isProximate +
                       Leave:FatherValence_s:isProximate:TimeDifference_s + Leave:FatherValenceVel:isProximate:TimeDifference_s +
                       Leave:MotherValence_s:MotherArousal_s + Leave:MotherValenceVel:MotherArousal_s +
                       Leave:FatherValence_s:isProximate:FatherArousal_s + Leave:FatherValenceVel:isProximate:FatherArousal_s +
                       Leave:FatherValence_s:isProximate:FatherArousal_s:TimeDifference_s + Leave:FatherValenceVel:isProximate:FatherArousal_s:TimeDifference_s | CoupleID / StudyDay))

p_coupled_dynamics <- c(
  prior(normal(0, 2), class = "b"),
  prior(normal(1, 2), class = "sd", resp = "FatherValenceAcc"),
  prior(normal(1, 2), class = "sd", resp = "FatherValences"),
  prior(normal(1, 2), class = "sd", resp = "FatherValenceVel"),
  prior(normal(1, 2), class = "sd", resp = "MotherValenceAcc"),
  prior(normal(1, 2), class = "sd", resp = "MotherValences"),
  prior(normal(1, 2), class = "sd", resp = "MotherValenceVel"),
  prior(normal(1, 2), class = "sigma", resp = "FatherValenceAcc"),
  prior(normal(1, 2), class = "sigma", resp = "FatherValences"),
  prior(normal(1, 2), class = "sigma", resp = "FatherValenceVel"),
  prior(normal(1, 2), class = "sigma", resp = "MotherValenceAcc"),
  prior(normal(1, 2), class = "sigma", resp = "MotherValences"),
  prior(normal(1, 2), class = "sigma", resp = "MotherValenceVel"),
  prior(lkj(2), class = "rescor")
)

CoupledOscillatorModel_time <- brm(
            f_coupled_dynamics_acc_time,
            data = d_multivariate,
            family = gaussian(),
            prior = p_coupled_dynamics, 
            sample_prior = "yes",
            chains = 2,
            cores = 64,
            warmup = 500,
            iter = 2000,
            control = list(adapt_delta = 0.90, max_treedepth = 10),
            backend = 'cmdstan')
```

## Extract Estimates for Supplementary Materials

```{r}
#CoupledOscillatorModel_time <- readRDS(here("FinalModels", "CoupledOscillatorModel_time.rds"))
tests_time <- extract_all_tests_time_model(CoupledOscillatorModel_time)
write.csv(tests_time, here("StatsForBodyText", "coupledoscillator_tests_time.csv"), row.names = FALSE)
```

# When Parents are Physically Distant
```{r, eval = FALSE}
f_coupled_dynamics_acc <- mvbf(
    FatherValence_s ~ 0 + Leave + (Leave | CoupleID),
    MotherValence_s ~ 0 + Leave + (Leave | CoupleID),
    FatherArousal_s ~ 0 + Leave + (Leave | CoupleID),
    MotherArousal_s ~ 0 + Leave + (Leave | CoupleID),
    FatherValenceVel ~ 0 + Leave + (Leave | CoupleID),
    MotherValenceVel ~ 0 + Leave + (Leave | CoupleID),
    FatherValenceAcc ~ 0 + Leave +
                       Leave:FatherValence_s + Leave:FatherValenceVel +
                       Leave:MotherValence_s  + Leave:MotherValenceVel +
                       Leave:FatherValence_s:FatherArousal_s + Leave:FatherValenceVel:FatherArousal_s +
                       Leave:MotherValence_s:MotherArousal_s + Leave:MotherValenceVel:MotherArousal_s +
                      (Leave:FatherValence_s + Leave:FatherValenceVel + Leave:MotherValence_s + Leave:MotherValenceVel + Leave:FatherValence_s:FatherArousal_s + Leave:FatherValenceVel:FatherArousal_s + Leave:MotherValence_s:MotherArousal_s + Leave:MotherValenceVel:MotherArousal_s | CoupleID),
    
    MotherValenceAcc ~ 0 + Leave +
                       Leave:MotherValence_s + Leave:MotherValenceVel +
                       Leave:FatherValence_s + Leave:FatherValenceVel +
                       Leave:MotherValence_s:MotherArousal_s + Leave:MotherValenceVel:MotherArousal_s +
                       Leave:FatherValence_s:FatherArousal_s + Leave:FatherValenceVel:FatherArousal_s +
                      (Leave:MotherValence_s + Leave:MotherValenceVel + Leave:FatherValence_s + Leave:FatherValenceVel + Leave:MotherValence_s:MotherArousal_s + Leave:MotherValenceVel:MotherArousal_s + Leave:FatherValence_s:FatherArousal_s + Leave:FatherValenceVel:FatherArousal_s | CoupleID))

CoupledOscillatorModel_farproximity <- brm(
            f_coupled_dynamics_acc,
            data = filter(d_multivariate, isProximate == 0),
            family = gaussian(),
            prior = p_coupled_dynamics, 
            sample_prior = "yes",
            chains = 2,
            cores = 64,
            warmup = 500,
            iter = 2000,
            control = list(adapt_delta = 0.99, max_treedepth = 15),
            backend = 'cmdstan')
```

## Extract Estimates for Supplementary Materials

```{r}
#CoupledOscillatorModel_farproximity <- readRDS(here("FinalModels", "CoupledOscillatorModel_farproximity.rds"))
coupledoscillator_tests_distant <- extract_all_tests_distant_parents(CoupledOscillatorModel_farproximity)
write.csv(coupledoscillator_tests_distant, here("StatsForBodyText", "coupledoscillator_tests_distant.csv"), row.names = FALSE)
```

## Plot of Relation between Arousal and Time of Day at Maternity and Paternity Leave
```{r}
ArousalPlotTimeOfDay <- ggplot() +
    geom_hline(yintercept = 0, linetype = "dashed", color = "black", alpha = 0.8) +
    # Add groups with different names in aes() to create legend entries
    geom_point(aes(x = TimeDecimals.x, y = FatherArousal_s, color = "Father"), 
               alpha = 0.4, data = d_multivariate, size = 1, 
               position = position_nudge(x = -.1)) +
    geom_point(aes(x = TimeDecimals.x, y = MotherArousal_s, color = "Mother"), 
               alpha = 0.4, data = d_multivariate, size = 1, 
               position = position_nudge(x = .1)) +
    geom_smooth(aes(x = TimeDecimals.x, y = MotherArousal_s, color = "Mother"), 
                data = d_multivariate) +
    geom_smooth(aes(x = TimeDecimals.x, y = FatherArousal_s, color = "Father"), 
                data = d_multivariate) +
    scale_x_continuous(breaks = seq(9, 22, 3), limits = c(9, 22)) +
    scale_y_continuous(breaks = seq(-2, 2, 1), limits = c(-2.5, 2.5)) +
    scale_color_manual(name = "Role",
                      values = c("Father" = "#21908CFF", "Mother" = "#D95F02")) +
    facet_wrap(~Leave) +
    ylab('Arousal') + xlab('Time of Day') + ggtitle('') +
    theme_minimal() +
    theme(
      plot.title = element_text(hjust = 0.5, size = 25, margin = margin(b = -20)),
      axis.text.x = element_text(size = 20, color = "black", margin = margin(t = -20)),
      axis.title.x = element_text(size = 20), 
      axis.text.y = element_text(size = 20, color = "black", margin = margin(t = -20)),
      axis.title.y = element_text(size = 20),
      axis.ticks = element_blank(),
      legend.position = "none",
      legend.text = element_text(size = 15),
      legend.title = element_blank(),
      legend.direction = "horizontal",
      axis.line.y = element_blank(),
      axis.line.x = element_blank(),
      legend.key = element_rect(fill = "transparent", color = NA),
      legend.background = element_rect(fill = "transparent", color = NA),
      strip.background = element_rect(
        color = "white", 
        fill = "white", 
        linewidth = 1.5, 
        linetype = "solid"
      ),
      strip.text.x = element_text(size = 20, color = "black"),
      plot.margin = margin(1, 1, 1, 1, "cm"),
      panel.grid.major = element_blank(),
      panel.grid.minor = element_blank()
    )

ValencePlotTimeOfDay <- ggplot() +
    geom_hline(yintercept = 0, linetype = "dashed", color = "black", alpha = 0.8) +
    geom_point(aes(x = TimeDecimals.x, y = FatherValence_s, color = "Father"), 
               alpha = 0.4, data = d_multivariate, size = 1, 
               position = position_nudge(x = -.1)) +
    geom_point(aes(x = TimeDecimals.x, y = MotherValence_s, color = "Mother"), 
               alpha = 0.4, data = d_multivariate, size = 1, 
               position = position_nudge(x = .1)) +
    geom_smooth(aes(x = TimeDecimals.x, y = MotherValence_s, color = "Mother"), 
                data = d_multivariate) +
    geom_smooth(aes(x = TimeDecimals.x, y = FatherValence_s, color = "Father"), 
                data = d_multivariate) +
    scale_x_continuous(breaks = seq(9, 22, 3), limits = c(9, 22)) +
    scale_y_continuous(breaks = seq(-2, 2, 1), limits = c(-2.5, 2.5)) +
    scale_color_manual(name = "Role",
                      values = c("Father" = "#21908CFF", "Mother" = "#D95F02")) +
    facet_wrap(~Leave) +
    ylab('Valence') + xlab('Time of Day') + ggtitle('') +
    theme_minimal() +
    theme(
      plot.title = element_text(hjust = 0.5, size = 25, margin = margin(b = -20)),
      axis.text.x = element_text(size = 20, color = "black", margin = margin(t = -20)),
      axis.title.x = element_text(size = 20), 
      axis.text.y = element_text(size = 20, color = "black", margin = margin(t = -20)),
      axis.title.y = element_text(size = 20),
      axis.ticks = element_blank(),
      legend.position = c(0.5, -0.2),
      legend.text = element_text(size = 15),
      legend.title = element_blank(),
      legend.direction = "horizontal",
      axis.line.y = element_blank(),
      axis.line.x = element_blank(),
      legend.key = element_rect(fill = "transparent", color = NA),
      legend.background = element_rect(fill = "transparent", color = NA),
      strip.background = element_rect(
        color = "white", 
        fill = "white", 
        linewidth = 1.5, 
        linetype = "solid"
      ),
      strip.text.x = element_text(size = 20, color = "black"),
      plot.margin = margin(1, 1, 1, 1, "cm"),
      panel.grid.major = element_blank(),
      panel.grid.minor = element_blank()
    )

PlotTimeOfDay <- plot_grid(ValencePlotTimeOfDay, ArousalPlotTimeOfDay, nrow = 2)

ggsave(plot = PlotTimeOfDay, file = here("FinalFigures", "PlotTimeOfDay.pdf"), height = 12, width = 9)
```

# Surrogate Couples

```{r, eval = FALSE}
# Identify couples who participated in both leave periods
both_periods_couples <- d_multivariate %>%
  group_by(CoupleID) %>%
  summarise(n_datapoints = n()) %>%
  filter(n_datapoints == 84) %>%
  pull(CoupleID)

# Filter the data to include only these couples
d_multivariate_both_periods <- d_multivariate %>%
  filter(CoupleID %in% both_periods_couples)

create_multiple_surrogates <- function(data, n_surrogates = 2, seed_val = 51) {
  # Get unique couples
  couples <- unique(data$CoupleID)
  
  # Function to create one surrogate dataset with a given seed
  create_one_surrogate <- function(seed_val) {
    set.seed(seed_val)
    shuffled_fathers <- sample(couples)
    
    # If any pair matches original, reshuffle
    while(any(couples == shuffled_fathers)) {
      shuffled_fathers <- sample(couples)
    }
    
    # Create surrogate dataset
    surrogate_data <- data
    
    # Initialize SurrogateFatherID column with original CoupleID
    surrogate_data$SurrogateFatherID <- surrogate_data$CoupleID
    
    # For each original couple
    for(i in seq_along(couples)) {
      orig_id <- couples[i]
      surr_id <- shuffled_fathers[i]
      
      # Get indices for the current couple
      orig_indices <- which(data$CoupleID == orig_id)
      surr_indices <- which(data$CoupleID == surr_id)
      
      # Replace father variables with surrogate father data
      surrogate_data$FatherValence_s[orig_indices] <- data$FatherValence_s[surr_indices]
      surrogate_data$FatherArousal_s[orig_indices] <- data$FatherArousal_s[surr_indices]
      surrogate_data$FatherValenceVel[orig_indices] <- data$FatherValenceVel[surr_indices]
      surrogate_data$FatherValenceAcc[orig_indices] <- data$FatherValenceAcc[surr_indices]
      
      # Assign the surrogate father ID to the current couple's rows
      surrogate_data$SurrogateFatherID[orig_indices] <- surr_id
    }
    
    # Add identifiers
    surrogate_data$SurrogateID <- paste0("S", seed_val)
    surrogate_data$isSurrogate <- TRUE
    
    return(surrogate_data)
  }
  
  # Create all surrogate datasets using different seeds
  surrogate_list <- map(1:n_surrogates, create_one_surrogate)
  
  # Add original data with appropriate identifiers
  original_data <- data %>%
    mutate(
      SurrogateID = "Original",
      isSurrogate = FALSE,
      SurrogateFatherID = CoupleID
    )
  
  # Combine all datasets
  all_data <- bind_rows(surrogate_list) %>%
    bind_rows(original_data)
  
  return(all_data)
}

surrogates_both_periods_S2 <- create_multiple_surrogates(d_multivariate_both_periods, n_surrogates = 1) %>%
  filter(isSurrogate == TRUE)

CoupledOscillatorModel_surrogate <- brm(
            f_coupled_dynamics_acc,
            data = surrogates_both_periods_S2,
            family = gaussian(),
            prior = p_coupled_dynamics, 
            sample_prior = "yes",
            chains = 2,
            cores = 64,
            warmup = 500,
            iter = 2000,
            control = list(adapt_delta = 0.90, max_treedepth = 10),
            backend = 'cmdstan')
```

## Extract Estimates for Supplementary Materials

```{r}
#CoupledOscillatorModel_surrogate <- readRDS(here("FinalModels", "CoupledOscillatorModel_surrogate.rds"))
tests_surrogate <- extract_all_tests_distant_parents(CoupledOscillatorModel_surrogate)
write.csv(tests_surrogate, here("StatsForBodyText", "coupledoscillator_tests_surrogate.csv"), row.names = FALSE)

surrogate_val_cor <- extract_residual_valence_correlation(
  model = CoupledOscillatorModel_surrogate, 
  TestDescription = "surrogate_val_cor"
)

far_prox_val_cor <- extract_residual_valence_correlation(
  model = CoupledOscillatorModel_farproximity, 
  TestDescription = "far_prox_val_cor"
)

surrogate_aro_cor <- extract_residual_arousal_correlation(
  model = CoupledOscillatorModel_surrogate, 
  TestDescription = "surrogate_aro_cor"
)

far_prox_aro_cor <- extract_residual_arousal_correlation(
  model = CoupledOscillatorModel_farproximity, 
  TestDescription = "far_prox_aro_cor"
)

phasespace_tests_control <- rbind(surrogate_val_cor, far_prox_val_cor, surrogate_aro_cor, far_prox_aro_cor)

write.csv(phasespace_tests_control, here("StatsForBodyText", "control_correlation_test.csv"))
```

# Get Close-Proximity Correlations for Each Leave Period

```{r, eval = FALSE}
f_coupled_dynamics_acc <- mvbf(
    FatherValence_s ~ 0 + (1 | CoupleID),
    MotherValence_s ~ 0 + (1 | CoupleID),
    
    FatherArousal_s ~ 0 + (1 | CoupleID),
    MotherArousal_s ~ 0 + (1 | CoupleID),
    
    FatherValenceVel ~ 0 + (1 | CoupleID),
    MotherValenceVel ~ 0 + (1 | CoupleID),
    
    FatherValenceAcc ~ 0 +
                       FatherValence_s + FatherValenceVel +
                       MotherValence_s:isProximate + MotherValenceVel:isProximate +
                       FatherValence_s:FatherArousal_s + FatherValenceVel:FatherArousal_s +
                       MotherValence_s:isProximate:MotherArousal_s + MotherValenceVel:isProximate:MotherArousal_s +
                      (FatherValence_s + FatherValenceVel + MotherValence_s:isProximate + MotherValenceVel:isProximate + FatherValence_s:FatherArousal_s + FatherValenceVel:FatherArousal_s + MotherValence_s:isProximate:MotherArousal_s + MotherValenceVel:isProximate:MotherArousal_s | CoupleID / StudyDay),
    
    MotherValenceAcc ~ 0 +
                       MotherValence_s + MotherValenceVel +
                       FatherValence_s:isProximate + FatherValenceVel:isProximate +
                       MotherValence_s:MotherArousal_s + MotherValenceVel:MotherArousal_s +
                       FatherValence_s:isProximate:FatherArousal_s + FatherValenceVel:isProximate:FatherArousal_s +
                      (MotherValence_s + MotherValenceVel + FatherValence_s:isProximate + FatherValenceVel:isProximate + MotherValence_s:MotherArousal_s + MotherValenceVel:MotherArousal_s + FatherValence_s:isProximate:FatherArousal_s + FatherValenceVel:isProximate:FatherArousal_s | CoupleID / StudyDay))

CoupledOscillatorModel_close_proximity_T1 <- brm(
            f_coupled_dynamics_acc,
            data = d_multivariate %>% filter(isProximate == 1) %>% filter(Leave == "Maternity Leave"),
            family = gaussian(),
            prior = p_coupled_dynamics, 
            sample_prior = "yes",
            chains = 2,
            cores = 64,
            warmup = 500,
            iter = 2000,
            control = list(adapt_delta = 0.90, max_treedepth = 10),
            backend = 'cmdstan')

CoupledOscillatorModel_close_proximity_T2 <- brm(
            f_coupled_dynamics_acc,
            data = d_multivariate %>% filter(isProximate == 1) %>% filter(Leave == "Paternity Leave"),
            family = gaussian(),
            prior = p_coupled_dynamics, 
            sample_prior = "yes",
            chains = 2,
            cores = 64,
            warmup = 1000,
            iter = 2500,
            control = list(adapt_delta = 0.999999, max_treedepth = 20),
            backend = 'cmdstan')
```

## Extract Estimates for Supplementary Materials
```{r}
CoupledOscillatorModel_close_proximity_T1 <- readRDS(here("FinalModels", "CoupledOscillatorModel_close_proximity_T1.rds"))
CoupledOscillatorModel_close_proximity_T2 <- readRDS(here("FinalModels", "CoupledOscillatorModel_close_proximity_T2.rds"))

T1_val_cor <- extract_residual_valence_correlation(
  model = CoupledOscillatorModel_close_proximity_T1, 
  TestDescription = "T1_val_cor"
)

T2_val_cor <- extract_residual_valence_correlation(
  model = CoupledOscillatorModel_close_proximity_T2, 
  TestDescription = "T2_val_cor"
)


T1_aro_cor <- extract_residual_arousal_correlation(
  model = CoupledOscillatorModel_close_proximity_T1, 
  TestDescription = "T1_aro_cor"
)

T2_aro_cor <- extract_residual_arousal_correlation(
  model = CoupledOscillatorModel_close_proximity_T2, 
  TestDescription = "T2_aro_cor"
)

phasespace_tests <- rbind(T1_val_cor, T2_val_cor,
                          T1_aro_cor, T2_aro_cor)


write.csv(phasespace_tests, here("StatsForBodyText", "correlation_tests.csv"))
```
