---
title: "simualate"
author: "Chris Cox"
date: "2025-11-28"
output: pdf_document
editor_options: 
  chunk_output_type: console
---

# Simulation Function for First Five Emotion Dynamics Patterns
```{r}
simulate_dyadic_emotions <- function(n_timepoints, 
                                   auto_corr_mother,
                                   auto_corr_father,
                                   cross_corr_m_to_f,
                                   cross_corr_f_to_m,
                                   correlation,
                                   sd_mother,
                                   sd_father) {
  
  # Create matrices to store the simulated values
  mother <- numeric(n_timepoints)
  father <- numeric(n_timepoints)
  
  # Generate correlated random noise
  mu <- c(0, 0)  # means
  sigma <- matrix(c(1, correlation, 
                   correlation, 1), 2, 2)  # correlation matrix
  
  # Generate multivariate normal random numbers
  random_effects <- MASS::mvrnorm(n_timepoints, mu = mu, Sigma = sigma)
  
  # Scale the random effects by the desired standard deviations
  noise_mother <- random_effects[, 1] * sd_mother
  noise_father <- random_effects[, 2] * sd_father
  
  # Initialize first time point
  mother[1] <- noise_mother[1]
  father[1] <- noise_father[1]
  
  # Generate time series using AR(1) process with cross-correlations
  for(t in 2:n_timepoints) {
    mother[t] <- auto_corr_mother * mother[t-1] + 
                 cross_corr_f_to_m * father[t-1] + 
                 noise_mother[t]
    
    father[t] <- auto_corr_father * father[t-1] + 
                 cross_corr_m_to_f * mother[t-1] + 
                 noise_father[t]
  }
  
  # Create output data frame
  result <- data.frame(
    time = 1:n_timepoints,
    mother = mother,
    father = father
  )
  
  return(result)
}
```

# First Pattern: Synchrony
```{r}
sim2.5 <- simulate_dyadic_emotions(
  n_timepoints = 18,
  auto_corr_mother = 0.3,
  auto_corr_father = 0.3,
  cross_corr_m_to_f = 0.4,  # Strong mutual influence
  cross_corr_f_to_m = 0.4,
  correlation = 0.95,        # High concurrent correlation
  sd_mother = 0.8,
  sd_father = 0.8
)

ValencePlot_1 <- ggplot(sim2.5, aes(x = time)) +
    geom_line(aes(y = mother, color = "Mother"), size = 1) +
    geom_line(aes(y = father, color = "Father"), size = 1) +
    geom_point(aes(y = mother, color = "Mother"), size = 3) +
    geom_point(aes(y = father, color = "Father"), size = 3) +
    #geom_vline(xintercept = seq(7, 38, by = 6), linetype = "dashed", alpha = 0.3) +
    scale_color_manual(values = c("#21908CFF", "#D95F02")) +
    geom_hline(yintercept = 0, color = "black", linetype = "dashed") +
    ylim(c(-2.5, 2.5)) +
    theme_minimal() +
    labs(title = "Synchrony",
         y = "Valence",
         x = "",
         color = "Partner") +
    theme(plot.title = element_text(hjust = 0.5, size = 20, face ="bold"),
          axis.text.x = element_blank(),
          axis.title.x = element_blank(),
          axis.text.y = element_blank(),
          axis.title.y = element_text(size = 20),
          legend.title = element_blank(),
          legend.text = element_text(size = 20),
          legend.key.size = unit(3, "lines"),
          legend.key = element_rect(fill = "transparent", color = NA),
          legend.background = element_rect(fill = "transparent", color = NA),
          legend.position = "none",
          strip.background = element_rect(color = "white", fill = "white", linewidth = 1.5, linetype = "solid"),
          panel.grid.minor.y = element_blank(),
          panel.grid.major.x = element_blank(),
          panel.grid.minor.x = element_blank())
ValencePlot_1
```

# Second Pattern: Co-Regulation
```{r}
sim3.5 <- simulate_dyadic_emotions(
  n_timepoints = 18,
  auto_corr_mother = 0.1,
  auto_corr_father = 0.1,
  cross_corr_m_to_f = 0.9,  # Strong mother-to-father influence
  cross_corr_f_to_m = 0.9,  # Strong father-to-mother influence
  correlation = 0.1,
  sd_mother = 0.5,
  sd_father = 0.5
)

ValencePlot_2 <- ggplot(sim3.5, aes(x = time)) +
    geom_line(aes(y = mother, color = "Mother"), size = 1) +
    geom_line(aes(y = father, color = "Father"), size = 1) +
    geom_point(aes(y = mother, color = "Mother"), size = 3) +
    geom_point(aes(y = father, color = "Father"), size = 3) +
    #geom_vline(xintercept = seq(7, 38, by = 6), linetype = "dashed", alpha = 0.3) +
    scale_color_manual(values = c("#21908CFF", "#D95F02")) +
    #scale_y_continuous(limits = c(0, 100), breaks = seq(0, 100, 20)) +
    #scale_x_continuous(limits = c(0, 42), breaks = seq(0, 42, 5)) +
    #scale_x_continuous(breaks = seq(1, 42, by = 6), labels = c("Day 1", "Day 2", "Day 3", "Day 4", "Day 5", "Day 6", "Day 7")) +
    geom_hline(yintercept = 0, color = "black", linetype = "dashed") +
    ylim(c(-2.5, 2.5)) +
    theme_minimal() +
    labs(title = "Co-Regulation",
         y = "Valence",
         x = "",
         color = "Partner") +
    theme(plot.title = element_text(hjust = 0.5, size = 20, face ="bold"),
          axis.text.x = element_blank(),
          axis.title.x = element_blank(),
          axis.text.y = element_blank(),
          axis.title.y = element_blank(),
          legend.title = element_blank(),
          legend.text = element_text(size = 20),
          legend.key.size = unit(3, "lines"),
          legend.key = element_rect(fill = "transparent", color = NA),
          legend.background = element_rect(fill = "transparent", color = NA),
          legend.position = "none",
          strip.background = element_rect(color = "white", fill = "white", linewidth = 1.5, linetype = "solid"),
          strip.text.x = element_text(size = 20, color = "black", face ="bold"),
          panel.grid.minor.y = element_blank(),
          panel.grid.major.x = element_blank(),
          panel.grid.minor.x = element_blank())
ValencePlot_2
```

# Third Pattern: Co-Dysregulation
```{r}
sim3 <- simulate_dyadic_emotions(
  n_timepoints = 18,
  auto_corr_mother = 0.5,    # High emotional inertia
  auto_corr_father = 0.5,    # High emotional inertia
  cross_corr_m_to_f = 0.75,   # Strong positive influence between partners
  cross_corr_f_to_m = 0.75,   # Same strong influence back
  correlation = 0.5,         # High concurrent correlation
  sd_mother = 0.6,          # High variability to allow for escalation
  sd_father = 0.6           # High variability to allow for escalation
)

ValencePlot_3 <- ggplot(sim3, aes(x = time)) +
    geom_line(aes(y = mother, color = "Mother"), size = 1) +
    geom_line(aes(y = father, color = "Father"), size = 1) +
    geom_point(aes(y = mother, color = "Mother"), size = 3) +
    geom_point(aes(y = father, color = "Father"), size = 3) +
    #geom_vline(xintercept = seq(7, 38, by = 6), linetype = "dashed", alpha = 0.3) +
    scale_color_manual(values = c("#21908CFF", "#D95F02")) +
    geom_hline(yintercept = 0, color = "black", linetype = "dashed") +
    #ylim(c(-5, 0)) +
    theme_minimal() +
    labs(title = "Co-Dysregulation",
         y = "Valence",
         x = "",
         color = "Partner") +
    theme(plot.title = element_text(hjust = 0.5, size = 20, face ="bold"),
          axis.text.x = element_blank(),
          axis.title.x = element_blank(),
          axis.text.y = element_blank(),
          axis.title.y = element_text(size = 20),
          legend.title = element_blank(),
          legend.text = element_text(size = 20),
          legend.key.size = unit(3, "lines"),
          legend.key = element_rect(fill = "transparent", color = NA),
          legend.background = element_rect(fill = "transparent", color = NA),
          legend.position = "none",
          strip.background = element_rect(color = "white", fill = "white", linewidth = 1.5, linetype = "solid"),
          strip.text.x = element_text(size = 20, color = "black", face ="bold"),
          panel.grid.minor.y = element_blank(),
          panel.grid.major.x = element_blank(),
          panel.grid.minor.x = element_blank())

ValencePlot_3
```

# Fourth Pattern: Leader-Follower
```{r}
sim5 <- simulate_dyadic_emotions(
  n_timepoints = 18,
  auto_corr_mother = 0.1,
  auto_corr_father = 0.1,
  cross_corr_m_to_f = 0.9,  # Strong mother-to-father influence
  cross_corr_f_to_m = 0.1,  # Weak father-to-mother influence
  correlation = 0.1,
  sd_mother = 1,
  sd_father = 1
)


ValencePlot_4 <- ggplot(sim5, aes(x = time)) +
    geom_line(aes(y = mother, color = "Mother"), size = 1) +
    geom_line(aes(y = father, color = "Father"), size = 1) +
    geom_point(aes(y = mother, color = "Mother"), size = 3) +
    geom_point(aes(y = father, color = "Father"), size = 3) +
    #geom_vline(xintercept = seq(7, 38, by = 6), linetype = "dashed", alpha = 0.3) +
    scale_color_manual(values = c("#21908CFF", "#D95F02")) +
    scale_color_manual(values = c("#21908CFF", "#D95F02")) +
    geom_hline(yintercept = 0, color = "black", linetype = "dashed") +
    ylim(c(-3, 3)) +
    theme_minimal() +
    labs(title = "Leader-Follower",
         y = "Valence",
         x = "",
         color = "Partner") +
  guides(color = guide_legend(nrow = 1)) +
    theme(plot.title = element_text(hjust = 0.5, size = 20, face = "bold"),
          axis.text.x = element_blank(),
          axis.title.x = element_text(size = 20, face = "bold"),
          axis.text.y = element_blank(),
          axis.title.y = element_blank(),
          strip.text = element_text(size = 20, face = "bold"),
          panel.grid.minor.y = element_blank(),
          panel.grid.major.x = element_blank(),
          panel.grid.minor.x = element_blank(),
          legend.title = element_blank(),
          legend.text = element_text(size = 20),
          legend.key.size = unit(2, "lines"),
          legend.key = element_rect(fill = "transparent", color = NA),
          legend.background = element_rect(fill = "transparent", color = NA),
          legend.position = c(0, 1.2),
          legend.box = "horizontal")
ValencePlot_4
```

# Fifth Pattern: Strong and Weak Self-Regulation
```{r}
# Create modified simulation function to focus on self-regulation
simulate_self_regulation <- function(n_timepoints, 
                                   auto_corr,
                                   position_force,
                                   sd_noise,
                                   startingpoint) {
  
  # Create vectors to store the simulated values
  position <- numeric(n_timepoints)
  velocity <- numeric(n_timepoints)
  
  # Initialize with perturbation from equilibrium
  position[1] <- startingpoint  # Start displaced from equilibrium
  velocity[1] <- 0  # Start with zero velocity
  
  # Generate time series using second-order dynamics
  for(t in 2:n_timepoints) {
    # Calculate acceleration based on position and velocity
    acceleration <- position_force * position[t-1] + auto_corr * velocity[t-1]
    
    # Update velocity using acceleration
    velocity[t] <- velocity[t-1] + acceleration + rnorm(1, 0, sd_noise)
    
    # Update position using velocity
    position[t] <- position[t-1] + velocity[t]
  }
  
  # Create output data frame
  result <- data.frame(
    time = 1:n_timepoints,
    emotion = position
  )
  
  return(result)
}

sim_weak_regulation <- simulate_self_regulation(
  n_timepoints = 18,
  auto_corr = -0.1,
  position_force = -0.23, 
  sd_noise = 0.3,
  startingpoint = 0.6
) %>%
  mutate(pattern = "Weak")

sim_strong_regulation <- simulate_self_regulation(
  n_timepoints = 18,
  auto_corr = -0.1,
  position_force = -1.1,
  sd_noise = 0.2,
  startingpoint = 0.3
) %>%
  mutate(pattern = "Strong")

SelfRegulationPlot <- ggplot() +
    geom_line(size = 1, color = "#e6ab02", data = sim_strong_regulation, aes(x = time, y = emotion)) +
    geom_point(size = 3, color = "#e6ab02", data = sim_strong_regulation, aes(x = time, y = emotion)) +
    geom_line(size = 1, color = "#7570b3", data = sim_weak_regulation, aes(x = time, y = emotion)) +
    geom_point(size = 3, color = "#7570b3", data = sim_weak_regulation, aes(x = time, y = emotion)) +
    geom_hline(yintercept = 0, color = "black", linetype = "dashed") +
    scale_y_continuous(limits = c(-2.5, 1.8), breaks = c(-1, 0, 1)) +
    theme_minimal() +
    labs(title = "Self-Regulation",
         y = "Valence",
         x = "Time") +
    theme(plot.title = element_text(hjust = 0.5, size = 20, face = "bold"),
          axis.text.x = element_blank(),
          axis.title.x = element_text(size = 20),
          axis.text.y = element_blank(),
          axis.title.y = element_text(size = 20),
          strip.text = element_text(size = 20, face = "bold"),
          panel.grid.minor.y = element_blank(),
          panel.grid.major.x = element_blank(),
          panel.grid.minor.x = element_blank())

SelfRegulationPlot
```

# Sixth Pattern: Strong and Weak Inertia
```{r}
sim_weak_inertia <- simulate_self_regulation(
  n_timepoints = 18,
  auto_corr = 0,
  position_force = -1, 
  sd_noise = 0.1,
  startingpoint = 0.5
) %>%
  mutate(pattern = "Weak")

sim_strong_inertia <- simulate_self_regulation(
  n_timepoints = 18,
  auto_corr = 0.1,
  position_force = -1,
  sd_noise = 0.1,
  startingpoint = 1
) %>%
  mutate(pattern = "Strong")

inertia_data <- rbind(
  sim_weak_inertia,
  sim_strong_inertia
)

InertiaPlot <- ggplot(data = inertia_data, aes(x = time, y = emotion, color = pattern)) +
    geom_line(size = 1) +
    geom_point(size = 3) +
    geom_hline(yintercept = 0, color = "black", linetype = "dashed") +
    scale_color_manual(values = c("Strong" = "#e6ab02", "Weak" = "#7570b3")) +
    scale_y_continuous(limits = c(-2.5, 1.8), breaks = c(-1, 0, 1)) +
    theme_minimal() +
    labs(title = "Inertia",
         y = "Emotion",
         x = "Time") +
  guides(color = guide_legend(nrow = 1)) +
    theme(plot.title = element_text(hjust = 0.5, size = 20, face = "bold"),
          axis.text.x = element_blank(),
          axis.title.x = element_text(size = 20),
          axis.text.y = element_blank(),
          axis.title.y = element_blank(),
          strip.text = element_text(size = 20, face = "bold"),
          panel.grid.minor.y = element_blank(),
          panel.grid.major.x = element_blank(),
          panel.grid.minor.x = element_blank(),
          legend.title = element_blank(),
          legend.text = element_text(size = 20),
          legend.key.size = unit(2, "lines"),
          legend.key = element_rect(fill = "transparent", color = NA),
          legend.background = element_rect(fill = "transparent", color = NA),
          legend.position = c(0, 0.1),
          legend.box = "horizontal")
InertiaPlot

SelfRegulation <- plot_grid(SelfRegulationPlot, InertiaPlot, nrow = 1, rel_widths = c(0.52, 0.48))
SelfRegulation
```

# Combine all plots in 
```{r}
SimulatedPatternsVisualAbstractFigure <- plot_grid(ValencePlot_1, ValencePlot_2, ValencePlot_3, ValencePlot_4, 
                                                   ncol = 2, rel_widths = c(0.52, 0.48))

SimulatedPatternsVisualAbstractFigure <- plot_grid(SimulatedPatternsVisualAbstractFigure, SelfRegulation, nrow = 2, rel_heights = c(0.66, 0.33))
SimulatedPatternsVisualAbstractFigure

ggsave(plot = SimulatedPatternsVisualAbstractFigure, filename = here("FinalFigures", "SimulatedPatternsVisualAbstractFigure.pdf"), height = 10, width = 12)
```

# Demonstration Plot for Position, Velocity, Acceleration
```{r}
damped_oscillator <- function(t, state, parameters) {
  with(as.list(c(state, parameters)), {
    dx_father <- v_father
    dv_father <- -k_father * x_father - b_father * v_father
    dx_mother <- v_mother
    dv_mother <- -k_mother * x_mother - b_mother * v_mother
    list(c(dx_father, dv_father, dx_mother, dv_mother))
  })
}

parameters <- c(
  k_father = 1.2,
  b_father = 0.8,
  k_mother = 0.5,
  b_mother = 0.3
)

initial_state <- c(
  x_father = -.5,
  v_father = 0,
  x_mother = -.5,
  v_mother = 0
)

times <- seq(0, 20, by = 0.1)
output <- ode(y = initial_state, times = times, func = damped_oscillator, parms = parameters)
output_df <- as.data.frame(output)
names(output_df) <- c("time", "father_emotion", "father_velocity", "mother_emotion", "mother_velocity")

# Calculate derivatives
output_df <- output_df %>%
  group_by(time) %>%
  # Velocity is already calculated in the ODE
  mutate(
    # Acceleration (second derivative)
    father_acceleration = -parameters["k_father"] * father_emotion - 
                         parameters["b_father"] * father_velocity,
    mother_acceleration = -parameters["k_mother"] * mother_emotion - 
                         parameters["b_mother"] * mother_velocity
  ) %>%
  ungroup()

# Add annotation at specific time point
t_point <- 4.6  # Choose point for annotation
point_data <- output_df %>% filter(near(time, t_point))

emotion_position <- ggplot(output_df) +
  geom_line(aes(x = time, y = father_emotion, color = "Father"), color = "#1B9E77", size = 1.5) +
  geom_line(aes(x = time, y = mother_emotion, color = "Mother"), color = "#D95F02", size = 1.5) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "gray50") +
  scale_color_manual(values = c("Father" = "#21908CFF", "Mother" = "#D95F02")) +
  geom_point(data = point_data,
             aes(x = time, y = father_emotion),
             color = "#21908CFF", size = 4) +
  geom_point(data = point_data,
             aes(x = time, y = mother_emotion),
             color = "#D95F02", size = 4) +
  scale_x_continuous(limits = c(0, 15), breaks = seq(0, 15, 3)) +
  scale_y_continuous(limits = c(-0.6, 0.6), breaks = seq(-1, 1, by = 0.5)) +
  labs(title = "Position", y = "Valence (z-scored)", x = "") +
  theme_minimal() +
    theme(plot.title = element_text(hjust = 0.5, size=20, color = "black"),
        axis.text.x = element_blank(),
        axis.title.x = element_text(size = 20, color = "black"),
        axis.text.y = element_text(size = 20, color = "black"),
        axis.title.y = element_text(size = 20, color = "black"),
        legend.title = element_blank(),
        legend.position = c(0.8, 0.2),
        legend.key.size = unit(3, "lines"),
        legend.text = element_text(size = 20),
        legend.background = element_rect(fill = "transparent", color = NA),
        legend.key = element_rect(fill = "transparent", color = NA),
        strip.background = element_rect(color="white", fill="white", linewidth=1.5, linetype="solid"),
        strip.text.x = element_text(size = 25, color = "black"),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank()
        )

emotion_velocity <- ggplot(output_df) +
  geom_line(aes(x = time, y = father_velocity, color = "Father"), color = "#1B9E77", size = 1.5) +
  geom_line(aes(x = time, y = mother_velocity, color = "Mother"), color = "#D95F02", size = 1.5) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "gray50") +
  scale_color_manual(values = c("Father" = "#21908CFF", "Mother" = "#D95F02")) +
  geom_point(data = point_data,
             aes(x = time, y = father_velocity),
             color = "#21908CFF", size = 4) +
  geom_point(data = point_data,
             aes(x = time, y = mother_velocity),
             color = "#D95F02", size = 4) +
  scale_x_continuous(limits = c(0, 15), breaks = seq(0, 15, 3)) +
  scale_y_continuous(limits = c(-0.6, 0.6), breaks = seq(-1, 1, by = 0.5)) +
  labs(title = "Velocity", y = "", x = "Time") +
  theme_minimal() +
    theme(plot.title = element_text(hjust = 0.5, size=20, color = "black"),
        axis.text.x = element_blank(),
        axis.title.x = element_text(size = 20, color = "black"),
        axis.text.y = element_text(size = 20, color = "black"),
        axis.title.y = element_text(size = 20, color = "black"),
        legend.title = element_blank(),
        legend.position = c(0.8, 0.2),
        legend.key.size = unit(3, "lines"),
        legend.text = element_text(size = 20),
        legend.background = element_rect(fill = "transparent", color = NA),
        legend.key = element_rect(fill = "transparent", color = NA),
        strip.background = element_rect(color="white", fill="white", linewidth=1.5, linetype="solid"),
        strip.text.x = element_text(size = 25, color = "black"),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank()
        )

emotion_acc <- ggplot(output_df) +
  geom_line(aes(x = time, y = father_acceleration, color = "Father"), color = "#1B9E77", size = 1.5) +
  geom_line(aes(x = time, y = mother_acceleration, color = "Mother"), color = "#D95F02", size = 1.5) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "gray50") +
  scale_color_manual(values = c("Father" = "#21908CFF", "Mother" = "#D95F02")) +
  geom_point(data = point_data,
             aes(x = time, y = father_acceleration),
             color = "#21908CFF", size = 4) +
  geom_point(data = point_data,
             aes(x = time, y = mother_acceleration),
             color = "#D95F02", size = 4) +
  scale_x_continuous(limits = c(0, 15), breaks = seq(0, 15, 3)) +
  scale_y_continuous(limits = c(-0.6, 0.6), breaks = seq(-1, 1, by = 0.5)) +
  labs(title = "Acceleration", y = "", x = "") +
  theme_minimal() +
    theme(plot.title = element_text(hjust = 0.5, size=20, color = "black"), 
        axis.text.x = element_blank(),
        axis.title.x = element_text(size = 20, color = "black"),
        axis.text.y = element_text(size = 20, color = "black"),
        axis.title.y = element_text(size = 20, color = "black"),
        legend.title = element_blank(),
        legend.position = c(0.8, 0.2),
        legend.key.size = unit(3, "lines"),
        legend.text = element_text(size = 20),
        legend.background = element_rect(fill = "transparent", color = NA),
        legend.key = element_rect(fill = "transparent", color = NA),
        strip.background = element_rect(color="white", fill="white", linewidth=1.5, linetype="solid"),
        strip.text.x = element_text(size = 25, color = "black"),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank()
        )

EmotionDynamicsDescription <- plot_grid(emotion_position, emotion_velocity, emotion_acc, nrow = 1)

#ggsave(plot = EmotionDynamicsDescription, filename = here("FinalFigures", "EmotionDynamicsDescription.pdf"), height = 6, width = 12)
```


# Strong, Medium and Weak Levels for Supplement

## First Pattern: Synchrony with Strong, Medium and Weak Levels
```{r}
# Strong Synchrony
sim_synchrony_strong <- simulate_dyadic_emotions(
  n_timepoints = 18,
  auto_corr_mother = 0.3,
  auto_corr_father = 0.3,
  cross_corr_m_to_f = 0.4,  
  cross_corr_f_to_m = 0.4,
  correlation = 0.95,
  sd_mother = 0.8,
  sd_father = 0.8
) %>%
  mutate(level = "Strong")

# Medium Synchrony
sim_synchrony_medium <- simulate_dyadic_emotions(
  n_timepoints = 18,
  auto_corr_mother = 0.3,
  auto_corr_father = 0.3,
  cross_corr_m_to_f = 0.4,  
  cross_corr_f_to_m = 0.4,
  correlation = 0.6,
  sd_mother = 0.8,
  sd_father = 0.8
) %>%
  mutate(level = "Medium")

# Weak Synchrony
sim_synchrony_weak <- simulate_dyadic_emotions(
  n_timepoints = 18,
  auto_corr_mother = 0.3,
  auto_corr_father = 0.3,
  cross_corr_m_to_f = 0.4,  
  cross_corr_f_to_m = 0.4,
  correlation = 0.25,
  sd_mother = 0.8,
  sd_father = 0.8
) %>%
  mutate(level = "Weak")

# Combine all data for plotting
synchrony_data <- bind_rows(
  sim_synchrony_strong,
  sim_synchrony_medium,
  sim_synchrony_weak
) %>%
  mutate(level = factor(level, levels = c("Strong", "Medium", "Weak"))) %>%
  pivot_longer(
    cols = c(mother, father),
    names_to = "partner",
    values_to = "emotion"
  ) %>%
  mutate(group = paste(level, partner, sep = "_"))

# Plot all synchrony levels
synchrony_plot <- ggplot(synchrony_data, aes(x = time, y = emotion, color = partner)) +
  geom_line(size = 1) +
  geom_point(size = 2) +
  scale_color_manual(values = c("mother" = "#21908CFF", "father" = "#D95F02"),
                     labels = c("mother" = "Mother", "father" = "Father")) +
  geom_hline(yintercept = 0, color = "black", linetype = "dashed") +
  ylim(c(-2.5, 2.5)) +
  facet_wrap(~ level, ncol = 3) +
  labs(title = "Synchrony",
       y = "Valence",
       x = "") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5, size = 20, face ="bold"),
          axis.text.x = element_blank(),
          axis.title.x = element_blank(),
          axis.text.y = element_blank(),
          axis.title.y = element_text(size = 20),
          legend.title = element_blank(),
          legend.text = element_text(size = 20),
          legend.key.size = unit(3, "lines"),
          legend.key = element_rect(fill = "transparent", color = NA),
          legend.background = element_rect(fill = "transparent", color = NA),
          legend.position = "none",
          strip.background = element_rect(color = "white", fill = "white", linewidth = 1.5, linetype = "solid"),
          strip.text.x = element_text(size = 20, color = "black", face ="bold"),
          panel.grid.minor.y = element_blank(),
          panel.grid.major.x = element_blank(),
          panel.grid.minor.x = element_blank())

synchrony_plot
```

## Second Pattern: Co-Regulation with Strong, Medium and Weak Levels
```{r}
# Strong Co-Regulation
sim_coreg_strong <- simulate_dyadic_emotions(
  n_timepoints = 18,
  auto_corr_mother = 0.1,
  auto_corr_father = 0.1,
  cross_corr_m_to_f = 0.9,
  cross_corr_f_to_m = 0.9,
  correlation = 0.1,
  sd_mother = 0.5,
  sd_father = 0.5
) %>%
  mutate(level = "Strong")

# Medium Co-Regulation
sim_coreg_medium <- simulate_dyadic_emotions(
  n_timepoints = 18,
  auto_corr_mother = 0.1,
  auto_corr_father = 0.1,
  cross_corr_m_to_f = 0.5,
  cross_corr_f_to_m = 0.5,
  correlation = 0.1,
  sd_mother = 0.5,
  sd_father = 0.5
) %>%
  mutate(level = "Medium")

# Weak Co-Regulation
sim_coreg_weak <- simulate_dyadic_emotions(
  n_timepoints = 18,
  auto_corr_mother = 0.1,
  auto_corr_father = 0.1,
  cross_corr_m_to_f = 0.2,
  cross_corr_f_to_m = 0.2,
  correlation = 0.1,
  sd_mother = 0.5,
  sd_father = 0.5
) %>%
  mutate(level = "Weak")

# Combine all data for plotting
coreg_data <- bind_rows(
  sim_coreg_strong,
  sim_coreg_medium,
  sim_coreg_weak
) %>%
  mutate(level = factor(level, levels = c("Strong", "Medium", "Weak"))) %>%
  pivot_longer(
    cols = c(mother, father),
    names_to = "partner",
    values_to = "emotion"
  ) %>%
  mutate(group = paste(level, partner, sep = "_"))

# Plot all co-regulation levels
coreg_plot <- ggplot(coreg_data, aes(x = time, y = emotion, color = partner)) +
  geom_line(size = 1) +
  geom_point(size = 2) +
  scale_color_manual(values = c("mother" = "#21908CFF", "father" = "#D95F02"),
                     labels = c("mother" = "Mother", "father" = "Father")) +
  geom_hline(yintercept = 0, color = "black", linetype = "dashed") +
  ylim(c(-2.5, 2.5)) +
  facet_wrap(~ level, ncol = 3) +
  labs(title = "Co-Regulation",
       y = "Valence",
       x = "") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5, size = 20, face ="bold"),
          axis.text.x = element_blank(),
          axis.title.x = element_blank(),
          axis.text.y = element_blank(),
          axis.title.y = element_text(size = 20),
          legend.title = element_blank(),
          legend.text = element_text(size = 20),
          legend.key.size = unit(3, "lines"),
          legend.key = element_rect(fill = "transparent", color = NA),
          legend.background = element_rect(fill = "transparent", color = NA),
          legend.position = "none",
          strip.background = element_rect(color = "white", fill = "white", linewidth = 1.5, linetype = "solid"),
          strip.text.x = element_text(size = 20, color = "black", face ="bold"),
          panel.grid.minor.y = element_blank(),
          panel.grid.major.x = element_blank(),
          panel.grid.minor.x = element_blank())

coreg_plot
```

## Third Pattern: Co-Dysregulation with Strong, Medium and Weak Levels
```{r}
# Strong Co-Dysregulation (high amplification)
sim_codysreg_strong <- simulate_dyadic_emotions(
  n_timepoints = 18,
  auto_corr_mother = 0.5,    
  auto_corr_father = 0.5,    
  cross_corr_m_to_f = 0.9,
  cross_corr_f_to_m = 0.9,   
  correlation = 0.5,        
  sd_mother = 0.6,          
  sd_father = 0.6           
) %>%
  mutate(level = "Strong")

# Medium Co-Dysregulation
sim_codysreg_medium <- simulate_dyadic_emotions(
  n_timepoints = 18,
  auto_corr_mother = 0.5,    
  auto_corr_father = 0.5,    
  cross_corr_m_to_f = 0.75,
  cross_corr_f_to_m = 0.75,   
  correlation = 0.5,        
  sd_mother = 0.6,          
  sd_father = 0.6           
) %>%
  mutate(level = "Medium")

# Weak Co-Dysregulation
sim_codysreg_weak <- simulate_dyadic_emotions(
  n_timepoints = 18,
  auto_corr_mother = 0.5,    
  auto_corr_father = 0.5,    
  cross_corr_m_to_f = 0.5,
  cross_corr_f_to_m = 0.5,   
  correlation = 0.5,        
  sd_mother = 0.6,          
  sd_father = 0.6           
) %>%
  mutate(level = "Weak")

# Combine all data for plotting
codysreg_data <- bind_rows(
  sim_codysreg_strong,
  sim_codysreg_medium,
  sim_codysreg_weak
) %>%
  # Reorder the factor levels before pivoting
  mutate(level = factor(level, levels = c("Strong", "Medium", "Weak"))) %>%
  pivot_longer(
    cols = c(mother, father),
    names_to = "partner",
    values_to = "emotion"
  ) %>%
  mutate(group = paste(level, partner, sep = "_"))

# Plot all co-dysregulation levels
codysreg_plot <- ggplot(codysreg_data, aes(x = time, y = emotion, color = partner)) +
  geom_line(size = 1) +
  geom_point(size = 2) +
  scale_color_manual(values = c("mother" = "#21908CFF", "father" = "#D95F02"),
                     labels = c("mother" = "Mother", "father" = "Father")) +
  geom_hline(yintercept = 0, color = "black", linetype = "dashed") +
  facet_wrap(~ level, ncol = 3) +
  labs(title = "Co-Dysregulation",
       y = "Valence",
       x = "Time") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5, size = 20, face ="bold"),
          axis.text.x = element_blank(),
          axis.title.x = element_blank(),
          axis.text.y = element_blank(),
          axis.title.y = element_text(size = 20),
          legend.title = element_blank(),
          legend.text = element_text(size = 20),
          legend.key.size = unit(3, "lines"),
          legend.key = element_rect(fill = "transparent", color = NA),
          legend.background = element_rect(fill = "transparent", color = NA),
          legend.position = "none",
          strip.background = element_rect(color = "white", fill = "white", linewidth = 1.5, linetype = "solid"),
          strip.text.x = element_text(size = 20, color = "black", face ="bold"),
          panel.grid.minor.y = element_blank(),
          panel.grid.major.x = element_blank(),
          panel.grid.minor.x = element_blank())

codysreg_plot
```

## Fourth Pattern: Leader-Follower with Strong, Medium and Weak Levels
```{r}
# Strong Leader-Follower (mother strongly influences father)
sim_leader_strong <- simulate_dyadic_emotions(
  n_timepoints = 18,
  auto_corr_mother = 0.1,
  auto_corr_father = 0.1,
  cross_corr_m_to_f = 0.9,
  cross_corr_f_to_m = 0.1,
  correlation = 0.1,
  sd_mother = 1,
  sd_father = 1
) %>%
  mutate(level = "Strong")

# Medium Leader-Follower
sim_leader_medium <- simulate_dyadic_emotions(
  n_timepoints = 18,
  auto_corr_mother = 0.1,
  auto_corr_father = 0.1,
  cross_corr_m_to_f = 0.6,
  cross_corr_f_to_m = 0.1,
  correlation = 0.1,
  sd_mother = 1,
  sd_father = 1
) %>%
  mutate(level = "Medium")

# Weak Leader-Follower
sim_leader_weak <- simulate_dyadic_emotions(
  n_timepoints = 18,
  auto_corr_mother = 0.1,
  auto_corr_father = 0.1,
  cross_corr_m_to_f = 0.3,
  cross_corr_f_to_m = 0.1,
  correlation = 0.1,
  sd_mother = 1,
  sd_father = 1
) %>%
  mutate(level = "Weak")

# Combine all data for plotting
leader_data <- bind_rows(
  sim_leader_strong,
  sim_leader_medium,
  sim_leader_weak
) %>%
  mutate(level = factor(level, levels = c("Strong", "Medium", "Weak"))) %>%
  pivot_longer(
    cols = c(mother, father),
    names_to = "partner",
    values_to = "emotion"
  ) %>%
  mutate(group = paste(level, partner, sep = "_"))

# Plot all leader-follower levels
leader_plot <- ggplot(leader_data, aes(x = time, y = emotion, color = partner)) +
  geom_line(size = 1) +
  geom_point(size = 2) +
  scale_color_manual(values = c("mother" = "#21908CFF", "father" = "#D95F02"),
                     labels = c("mother" = "Mother", "father" = "Father")) +
  geom_hline(yintercept = 0, color = "black", linetype = "dashed") +
  ylim(c(-3, 3)) +
  facet_wrap(~ level, ncol = 3) +
  labs(title = "Leader-Follower",
       y = "Valence",
       x = "Time") +
  theme_minimal() +
  guides(color = guide_legend(nrow = 1)) +
  theme(plot.title = element_text(hjust = 0.5, size = 20, face ="bold"),
          axis.text.x = element_blank(),
          axis.title.x = element_text(size = 20),
          axis.text.y = element_blank(),
          axis.title.y = element_text(size = 20),
          legend.title = element_blank(),
          legend.text = element_text(size = 20),
          legend.key.size = unit(3, "lines"),
          legend.key = element_rect(fill = "transparent", color = NA),
          legend.background = element_rect(fill = "transparent", color = NA),
          legend.position = c(0.8, 1.5),
          strip.background = element_rect(color = "white", fill = "white", linewidth = 1.5, linetype = "solid"),
          strip.text.x = element_text(size = 20, color = "black", face ="bold"),
          panel.grid.minor.y = element_blank(),
          panel.grid.major.x = element_blank(),
          panel.grid.minor.x = element_blank())
leader_plot
```


```{r}
final_plot <- plot_grid(
  synchrony_plot, coreg_plot, 
  codysreg_plot, leader_plot, 
  nrow = 4, 
  align = 'vh'
)

#ggsave(plot = final_plot, file = here("FinalFigures", "emotion_dynamics_strongweakpatterns.png"), width = 12, height = 14)
```

